Objectification in Action: Self- and Other-Objectification in Same-gender Interactions

### 2020-11-10 Final_Manuscript.tex
measures table

\subsection{Post interaction Measures}\label{post-interaction-measures}

@@

The following measures were collected in the order they are presented
following the interaction. Correlations and descriptive statistics of
all study variables appear in Table~\ref{tab:corrtable}.

\subsubsection{Cognitive Performance}\label{cognitive-performance}

Trigrams from the Remote Associates Task (McFarlin \& Blascovich, 1984)
were utilized to assess cognitive performance after the interaction. Ten
items were selected and presented to participants. For example, the
correct answer for the trigram \enquote{Quack: Pond: Waddle} would be
\enquote{Duck}. Participants are limited to 30 seconds to provide their
answers. For every correct answer, 1 point is given. The mean score was
5.03 (SD = 2.29). Cognitive performance was measured first directly
after the interaction in order to measure potential immediate detriments
to performance (Garcia et al., 2016).

\subsubsection{State
Other-Objectification}\label{state-other-objectification}

To measure the participant's objectification of their partner in the
interaction, participants were asked a series of questions about the
frequency of thoughts during the interaction in relation to multiple
characteristics of their partner (Garcia et al., 2016). Questions
included aspects of their partner's internal traits such as personality,
friends, family, and extracurricular interests, as well as external
traits such as body, appearance, clothing, and body parts. All questions
were rated on a scale from 1 (not at all) to 7 (constantly).
Objectification was measured by taking the difference between the
average frequency of thought about their partner's external traits
(\(\alpha\) = 0.79 for Sample 1, \(\alpha\) = 0.79 for Sample 2) and
frequency of thought about their partner's internal traits (\(\alpha\) =
0.79 for Sample 1, \(\alpha\) = 0.76 for Sample 2). A positive score in
this scale would indicate that the participant thought about their
partner's external traits more than the partner's internal traits, and a
negative score would indicate the opposite.

@@

\subsubsection{Interaction Authenticity}\label{interaction-authenticity}

To assess the magnitude to which individuals felt comfortable in the
interaction and perceived the interaction to be authentic, we asked
participants to rate the extent to which they felt comfortable, happy,
friendly, warm, easygoing, sincere, and authentic on a scale ranging
from 1 (not at all) to 7 (very much), much like (Garcia et al., 2016).
Participants were additionally asked to rate their interaction partner's
authenticity as well as their own: \enquote{Do you think your partner
was authentic during your interaction?} and \enquote{Were you authentic
during your interaction?} These questions were ranked on a scale from 1
(not at all) to 7 (very much). These were combined to form the
authenticity scale (\(\alpha\) = 0.91 for Sample 1, \(\alpha\) = 0.91
for Sample 2).

\subsubsection{State
Self-Objectification}\label{state-self-objectification}

To assess state self-objectification, we used an average of two items
from Saguy et al. (2010) that were also used in Garcia et al. (2016).
Participants were asked to rank how much they agreed with the following
statements: \enquote{During the interaction I felt more like a body than
a full self} and \enquote{I felt more like a body than as a real person
in the interaction}. Originally, Saguy et al. (2010) used 3 items, but
in both samples the reliability of the scale was higher once the third
item was removed, so we chose to only use the first two for our measure
of SSO, leaving us with a reliable scale (\(\alpha\) = 0.84 for Sample
1, and \(\alpha\) = 0.85 for Sample 2.)

\subsubsection{Relationship Agency}\label{relationship-agency}

A scale was used from Garcia et al. (2016) to assess how much agency an
individual believes they would possess in future romantic relationships.
Participants were asked how likely it was that they would do the
following: \enquote{ask someone out on a date,} \enquote{open the door
for your date,} \enquote{pay for a date,} \enquote{ask your
boyfriend/girlfriend to marry you,} \enquote{initiate sex with your
girlfriend/boyfriend,} \enquote{initiate condom use during sex,}
\enquote{surprise your boyfriend/ girlfriend with a gift,} and
\enquote{ask your girlfriend/boyfriend to move with you to a new place.}
Responses were measured on a scale ranging from 1 (not at all likely) to
7 (extremely likely). The scale originally had 9 items, but the 9th item
had low correlations with the remaining items, ranging from .02 to .30
for the first sample, and .04 to .30 for the second sample. Therefore,
this item was removed. As a result, the 8-item scale had moderately high
reliability for both samples (\(\alpha\) = 0.72 for Sample 1,
(\(\alpha\) = 0.74 for Sample 2).

\subsubsection{Career Aspirations}\label{career-aspirations}

To conceptualize participants' career aspirations after the interaction,
we used the 10-item adaptation of Gray and OBrien (2007)'s Career
Aspiration Scale which asked participants to consider how true 10
statements were in regard to their future careers on a scale from 0 (not
at all true of me) to 4 (very true of me). Items include \enquote{I hope
to become a leader in my career field} and \enquote{I hope to move up
through any organization or business I work in.} Items were fairly
reliable (\(\alpha\) = 0.73 for Sample 1, \(\alpha\) = 0.80 for Sample
2).

\subsubsection{Trait
Self-Objectification}\label{trait-self-objectification}

Trait self-objectification (TSO) was assessed using the
Self-Objectification Questionnaire (B. L. Fredrickson et al., 1998; S.
Noll \& Fredrickson, 1998), which evaluates the extent to which
individuals view their bodies in observable versus non-observable ways.
The questionnaire asked participants to rank order both appearance and
functional aspects of their bodies, from 1 (least important) to 10 (most
important), with respect to physical self-concepts. Of the ten body
attributes, five of the items were appearance-based (weight, sex appeal,
physical attractiveness, firm/sculpted muscles and body measurements),
and five of the items were competence-based (strength, physical
coordination, energy level, health and physical fitness). Difference
scores were computed by subtracting the sum of the 5 functional
aspects/competence attributes (e.g., health, strength) from the sum of
the 5 physical self-concepts/appearance attributes (e.g., physical
attractiveness, weight), and all measures were multiplied by -1 so that
positive scores indicated greater TSO.

@@

In addition to being relatively small and combined across two
institutions, another limitation of the current study sample is that the
it was comprised of Western women only. Being that self-objectification
has been found to be most prevalent in Western culture (Loughnan et al.,
2015) research on objectification conducted outside of Western or
Westernized countries has been scarce (Moradi \& Huang, 2008), although
more recent work has examined objectification from a cross-cultural
framework (Loughnan et al., 2015, Wollast et al. (2020)). Since
\enquote{bodies exist within social and cultural contexts, and hence are
also constructed through sociocultural practices and discourses}
(Fredrickson \& Roberts, 1997, p. 174), it is important to consider how
diverse social identities within unique cultural contexts may inform
sexual objectification phenomenon to test the cross-cultural
applicability of theoretical frameworks (Loughnan et al., 2015).

Further, sexualizing experiences and self-objectification are thought to
begin a very young age, and thus, researchers have only recently begun
to examine such experiences among children (e.g., Holland \& Haslam,
2016; Jongenelis, Byrne, \& Pettigrew, 2014). Considering the fact that
the average mean age of the investigated participants of this current
study was 18.85 years, research among younger and older individuals is
needed, especially because the processes of self-objectification may
change over time (Fredrickson \& Roberts, 1997). It may be valuable to
question the extent to which children, adolescents, or emerging adults
of different races or ethnicities are exposed to varied amounts of
sexualizing content. Further, more longitudinal studies investigating
the developmental experience of interpersonal objectification and
self-objectification are needed.

@@

The current sample contained a mixture of heterosexual and
non-heterosexual women, but all participants were asked to think about
and evaluate their partner as a potential dating/romantic partner. We
think that heterosexual women are able to do this with other women as
their target---and there is evidence that they might do this readily
(Puvia \& Vaes, 2013; Strelan \& Hargreaves, 2005)---perhaps they may be
even more apt to activate social comparison processes (Festinger, 1954)
than women who are sexually attracted to other women (non-heterosexual
women). Since both men and women are socialized in a culture that
sexually objectifies women, both men and women may come to internalize
this socialization and sexually objectify women. Indeed, recent research
has found that the more women self-objectify, the more they objectify
other women (Loughnan et al., 2015; Puvia \& Vaes, 2013), although not
to the degree exhibited by men; that is, men were found to objectify
women significantly more than women objectify other women (Strelan \&
Hargreaves, 2005). This differential psychological process between women
with differing sexualities might have served to dampen our ability to
detect relationships.

Previous research has found that when compared to heterosexual women,
lesbian women report less concern with physical appearance (Siever,
1994; Strong, Williamson, Netemeyer, \& Geer, 2000), lower body
surveillance (Hill \& Fischer, 2008), and lower self-objectification
(Noffsinger-Frazier, 2004). However none of these studies examined the
relationship between self-objectification and experiences of sexual
objectification. Thus, it is unclear whether lesbian women indeed
experience similar levels of cultural sexual objectification but
internalize them less than heterosexual women do. Consistent with
previous research, Hill and Fischer (2008) determined that lesbians
exhibited less physical appearance concerns compared to heterosexual
women, however there was no evidence found that lesbian women
self-objectify less than heterosexual women and they did not find that
sexual orientation moderates the relationship between sexual
objectification and self-objectification. This evidence sits in contrast
to older theoretical literature that suggests that lesbians internalize
cultural sexual objectification less than do heterosexual women (Brown,
1987; Pitman, 1999; Rothblum, 1994; Siever, 1994).

@@

The results of this study demonstrate the complex and ambivalent nature
of sexual objectification and additionally highlight the psychological
and social consequences of such objectification processes on women's
social relationships and well-being. These results are quite useful for
promoting mental health, the creation and maintenance of early action
programs for girls and young women, and for scholars and practitioners
to work intentionally to provide the tools necessary to circumvent or
mitigate negative effects on self-objectification to combat such
experiences.

@@

\hypertarget{ref-brown1987lesbians}{}
Brown, L. S. (1987). Lesbians, weight, and eating: New analyses and
perspectives. \emph{Lesbian Psychologies: Explorations and Challenges},
294--309.

@@

@@

\hypertarget{ref-grayobrien2007}{}
Gray, M. P., \& OBrien, K. M. (2007). Advancing the assessment of
women's career choices: The career aspiration scale. \emph{Journal of
Career Assessment - J CAREER ASSESSMENT}, \emph{15}, 317--337.
doi:\href{https://doi.org/10.1177/1069072707301211}{10.1177/1069072707301211}

@@

@@

\hypertarget{ref-noffsinger2004objectification}{}
Noffsinger-Frazier, N. A. (2004). \emph{Objectification theory and
disordered eating: The impact of feminist identification,
internalization of sociocultural standards of appearance, and sexual
orientation} (PhD thesis). The University of Memphis.

\hypertarget{ref-nollfredrickson1998}{}
Noll, S., \& Fredrickson, B. (1998). A mediational model linking
self-objectification, body shame, and disordered eating.
\emph{Psychology of Women Quarterly}, \emph{22}, 623--636.
doi:\href{https://doi.org/10.1111/j.1471-6402.1998.tb00181.x}{10.1111/j.1471-6402.1998.tb00181.x}

@@

\hypertarget{ref-pitman1999body}{}
Pitman, G. E. (1999). Body image, compulsory heterosexuality, and
internalized homophobia. \emph{Journal of Lesbian Studies}, \emph{3}(4),
129--139.

@@

\hypertarget{ref-rothblum1994lesbians}{}
Rothblum, E. (1994). Lesbians and physical appearance which model
applies? \emph{Lesbian and Gay Psychology: Theory, Research, and
Clinical Applications}, \emph{1}, 84.

@@

\hypertarget{ref-siever1994sexual}{}
Siever, M. D. (1994). Sexual orientation and gender as factors in
socioculturally acquired vulnerability to body dissatisfaction and
eating disorders. \emph{Journal of Consulting and Clinical Psychology},
\emph{62}(2), 252.

@@

\hypertarget{ref-strong2000eating}{}
Strong, S. M., Williamson, D. A., Netemeyer, R. G., \& Geer, J. H.
(2000). Eating disorder symptoms and concerns about body differ as a
function of gender and sexual orientation. \emph{Journal of Social and
Clinical Psychology}, \emph{19}(2), 240--255.

>>>>

\subsection{Measures}\label{measures}

@@

\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\caption{\label{tab:measures}Measures}
\begin{tabular}{lllll}
\toprule
 & \multicolumn{1}{c}{Original Source} & \multicolumn{1}{c}{Example Item} & \multicolumn{1}{c}{Reliability Sample 1} & \multicolumn{1}{c}{Reliability Sample 2}\\
\midrule
Cognitive performance & @mcfarlin1984remote & The correct answer for the trigram "Quack: Pond: Waddle" would be "Duck" & Not avaliable & Not avaliable\\
Objectification of partner & @garcia2016objectification & "During the interaction, how often did you think about your partner’s personality?" (...friends, family, and extracurricular interests, body, appearance, clothing, and body parts) on a scale from 1 (not at all) to 7 (constantly). A difference between the average internal and exxternal traits is taken & external:  0.79 ; internal:  0.71 & external:  0.76 ; internal:  0.79\\
Authenticity of interaction & @garcia2016objectification & "During the interaction, to what extent did you feel comfortable?" (...happy, friendly, warm, easygoing, sincere, and authentic) on a scale ranging from 1 (not at all) to 7 (very much) & 0.91 & 0.91\\
State self-objectification & @saguyetal2010 & "During the interaction I felt more like a body than a full self" & 0.84 & 0.85\\
Relationship agency & @garcia2016objectification & "ask someone out on a date" Responses were measured on a scale ranging from 1 (not at all likely) to 7 (extremely likely) & 0.72 & 0.74\\
Career aspirations & @grayobrien2007 & "I hope to become a leader in my career field" & 0.73 & 0.80\\
Trait self objectification (TSO) & @nollfredrickson1998 & Participants rank order aspects of their bodies from 1 (least important) to 10 (most important): weight, sex appeal, physical attractiveness, firm/sculpted muscles and body measurements, strength, physical coordination, energy level, health and physical fitness. Difference scores were computed by subtracting the sum of the 5 functional aspects/competence attributes (e.g., health, strength) from the sum of the 5 physical self-concepts/appearance attributes (e.g., physical attractiveness, weight), and all measures were multiplied by -1 so that positive scores indicated greater TSO. & appearance:  0.44 ; competence:  0.43 & appearance:  0.51 ; competence:  0.30\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\end{table}

The same set of measures given in Garcia et al. (2016) were given to
participants in the current study. For more details about each of these
measures please see that previous paper. To summarize, the list of
measures in the order given to participants appears in
Table~\ref{tab:measures} along side the orginal source for the scale, an
example item, and the reliability in the current sample. Correlations
and descriptive statistics of all study variables appear in
Table~\ref{tab:corrtable}.

Cognitive performance with the Remote Associates Task's trigrams
(McFarlin \& Blascovich, 1984) was given first, to capture any
attentional changes that may have been affected by the interaction. Next
participants were given a set of measure about their thoughts and
feelings during the interaction that included State-Other
Objectification, Interaction Authenticity, and State
Self-Objectification. Following these measures participants completed
the Relationship Agency Scale and the Career Aspirations Scale. Lastly,
as was the case in Garcia et al. (2016), Trait Self-Objectification was
given last, as not to arouse suspicious at the start of the study that
we were actually focusing on objectification and to give enough time
that immediate effects of the interaction would have worn off.

@@

Howbody used 3 items, but in both samples the reliability of the scale
was higher once the third item was removed, so we chose to only use the
first two for our measure of SSO, leaving us with a reliable scale (see
Table~\ref{tab:measures}).

Relationship agency The scale originally had 9 items, but the 9th item
had low correlations with the remaining items, ranging from .02 to .30
for the first sample, and .04 to .30 for the second sample. Therefore,
this item was removed. As a result, the 8-item scale had moderately high
reliability for both samples (see Table~\ref{tab:measures}).

@@

In addition to being relatively small and combined across two
institutions, another limitation of the current study sample is that the
it was comprised of Western women only. Being that self-objectification
has been found to be most prevalent in Western culture (Loughnan et al.,
2015) research on objectification {[}{[}{[}conducted outside of Western
or Westernized countries{]}{]}{]} has been scarce (Moradi \& Huang,
2008), although more recent work has examined objectification from a
cross-cultural framework (Loughnan et al., 2015, Wollast et al. (2020)).
{[}{[}{[}Since \enquote{bodies exist within social and cultural
contexts, and hence are also constructed through sociocultural practices
and discourses}{]}{]}{]} (Fredrickson \& Roberts, 1997, p. 174), it is
important to consider how diverse social identities within unique
cultural contexts may inform sexual objectification phenomenon
{[}{[}{[}to test the cross-cultural applicability of theoretical
frameworks{]}{]}{]} (Loughnan et al., 2015).

Further, sexualizing experiences and self-objectification are thought to
begin a very young age, and thus, researchers have only recently begun
to examine such experiences among children (e.g., Holland \& Haslam,
2016; Jongenelis, Byrne, \& Pettigrew, 2014). {[}{[}Considering the fact
that the average mean age of the investigated participants of this
current study was 18.85 years, research among younger and older
individuals is needed{]}{]}{]}, especially because the processes of
self-objectification may change over time (Fredrickson \& Roberts,
1997). {[}{[}{[}It may be valuable to question the extent to which
children, adolescents, or emerging adults of different races or
ethnicities are exposed to varied amounts of sexualizing
content.{]}{]}{]} Further, more longitudinal studies investigating the
developmental experience of interpersonal objectification and
self-objectification are needed.

@@

The current sample contained a mixture of heterosexual and
non-heterosexual women, but all participants were asked to think about
and evaluate their partner as a potential dating/romantic partner. We
think that heterosexual women are able to do this with other women as
their target---and there is evidence that they might do this readily
(Puvia \& Vaes, 2013; Strelan \& Hargreaves, 2005)---perhaps they may be
even more apt to activate social comparison processes (Festinger, 1954)
than women who are sexually attracted to other women (non-heterosexual
women). {[}{[}{[}Indeed, recent research has found that the more women
self-objectify, the more they objectify other women{]}{]}{]} (Loughnan
et al., 2015; Puvia \& Vaes, 2013), {[}{[}{[}although not to the degree
exhibited by men; that is, men were found to objectify women
significantly more than women objectify other women{]}{]}{]} (Strelan \&
Hargreaves, 2005). This differential psychological process between women
with differing sexualities might have served to dampen our ability to
detect relationships.

\subsubsection{Sexism and Sexual
Discrimination}\label{sexism-and-sexual-discrimination}

Sexism has been found to be related to the objectification of women, but
not men (Cikara, Eberhardt, \& Fiske, 2011; Harsey \& Zurbriggen, 2020),
and in addition, TSO has been found to be associated with less negative
attitudes towards sexual harassment (Bernard, Legrand, \& Klein, 2018).
Both of these effects are present for both men and women. Empirical
evidence illustrates how women continue to be objects of interpersonal
discrimination and experience daily sexist hassles (Swim, Hyers, Cohen,
\& Ferguson, 2001). One form of interpersonal discrimination women face
is the process by which their whole being is viewed as a collection of
sexualized body parts valued predominantly for commodification, a
phenomena termed sexual objectification (Bartky, 1990). Sexual
objectification of women has also been linked to trauma (Miles-McLean et
al., 2015). Sexual objectification occurs in a myriad of ways in women's
lives and is modified by unique combinations of race, ethnicity,
sexuality, age, and class (Fredrickson \& Roberts, 1997). Amid such
heterogeneity though, \enquote{having a reproductively mature female
body} proposed by Fredrickson and Roberts (1997) is likely to create a
shared vulnerability to sexual objectification and a variety of shared
negative experiences as a result.

@@

The results of this study \ldots{} and additionally highlight the
psychological and social consequences of such objectification processes
on women's social relationships and well-being. These results are quite
useful for promoting mental health, the creation and maintenance of
early action programs for girls and young women, and for scholars and
practitioners to work intentionally to provide the tools necessary to
circumvent or mitigate negative effects on self-objectification to
combat such experiences.

@@

\hypertarget{ref-bernard2018bodies}{}
Bernard, P., Legrand, S., \& Klein, O. (2018). From bodies to blame:
Exposure to sexually objectifying media increases tolerance toward
sexual harassment. \emph{Psychology of Popular Media Culture},
\emph{7}(2), 99.

@@

\hypertarget{ref-cikara2011agents}{}
Cikara, M., Eberhardt, J. L., \& Fiske, S. T. (2011). From agents to
objects: Sexist attitudes and neural responses to sexualized targets.
\emph{Journal of Cognitive Neuroscience}, \emph{23}(3), 540--551.

@@

@@

\hypertarget{ref-miles2015stop}{}
Miles-McLean, H., Liss, M., Erchull, M. J., Robertson, C. M., Hagerman,
C., Gnoleba, M. A., \& Papp, L. J. (2015). ``Stop looking at me!''
interpersonal sexual objectification as a source of insidious trauma.
\emph{Psychology of Women Quarterly}, \emph{39}(3), 363--374.

@@

@@

@@

@@

@@

\hypertarget{ref-swimetal}{}
Swim, J. K., Hyers, L. L., Cohen, L. L., \& Ferguson, M. J. (2001).
Everyday sexism: Evidence for its incidence, nature, and psychological
impact from three daily diary studies. \emph{Journal of Social Issues},
\emph{57}(1), 31--53.
doi:\href{https://doi.org/10.1111/0022-4537.00200}{10.1111/0022-4537.00200}
